The DAO Hack

### 2019-01-07 paper/paper.tex
added cites
added fallout and conclusion

A benefit of the blockchain is that no single party can change past transactions.
Thus, once something is in the blockchain, it is very hard to get rid of it. \cite{blockchainKai}

@@

@@

Ethereum is one type of cryptocurrency.
The cash is calles ether.
Its main difference to others like Bitcoin is that it can run a complex programming language called Solidity.
With Solidity, you can create loops, functions and recursion which then are run on the ethereum blockchain.
This has huge advantages as one can create powerful smart contracts and bind the payment to those.
Only with such a powerful language the DAO could be created. \cite{eth}

@@

An investment fund is used to invest money. A investor can give money to a fond which will in turn invest the money by buying shares and other investment opportunities. Usually a fond will have a big diversification between their investments, often bigger than any single entity could achieve. A fond usually has some guidlines on how to invest the money based on risk and other factors. Those factors mostly differ a lot between fonds. As an example there are fonds that focus on the swiss market, or on the telecomunication market world wide.

The fond usually has someone in charge of the investments called fonds manager. He has the power to invest the money. Additionally to shares fonds also invest in proposals. A proposal is generated by companies where they offer investment opportunities for a specific project or program. The fond manager then bids on the proposals he thinks will be profitable.

@@

In May 2016 when the DAO was launched on the ethereum blockchain it was funded with 12.7M ether.
As one ether was worth about 10-12 Dollars at that time, this corresponds to about 150 million dollars.

@@

The split function contains a function called withdrawRewardFor which calls rewardAccount.payOut which in turn calls \textunderscore recipient.call.value. The first two functions are defined in the DAO code and as such can't be changed at will. The call to recipient however is in the attackers hands. He can do there whatever he desires. In code this looks like that: \cite{deconstructingDaoAttack}

@@

@@

@@

>>>>

A benefit of the blockchain is that no single party can change past transactions.
Thus, if a transaction is burried in the blockchain deep enough it is very hard to remove it. Which is different in classical currency where a bank can always change what has happened by accepting or denying transactions and even changing past transactions is possible. \cite{blockchainKai}

@@

\subsection{Soft Fork}

A soft fork on the other hand is used often when security is increased. Important for a soft fork is that fewer blocks are valid. With that in mind any newly created block by the new version of the software is still valid with the old version, but one with the old version might be invalid for the newer one. With this not everyone has to switch at the same time and the same cryptocurrency can continue to live. Once the majority has adopted the new and stronger rules can be activated and the blocks generated with the old version will be ignored. This way everybody is incentivised to switch to the new software version so that their blocks are valid. \cite{blockchainKai}

@@

Ethereum is one type of cryptocurrency.
The cash is calles ether.
Its main difference to others like Bitcoin is that it can run a complex programming language called Solidity.
With Solidity, you can create loops, functions and recursion which then are run on the ethereum blockchain.
This has huge advantages as one can create powerful smart contracts and bind the payment to those.
Only with such a powerful language the DAO could be created. 
A smart contract that runs on ethereum can not be stopped as it is run on the blockchain. \cite{eth}

@@

An investment fund is used to invest money. A investor can give money to a fond which will in turn invest the money by buying shares and other investment opportunities. Usually a fond will have a big diversification between their investments, often bigger than any single entity could achieve. A fond usually has some guidlines on how to invest the money based on risk and other factors. Those factors mostly differ a lot between fonds. As an example there are fonds that focus on the swiss market, or on the telecomunication market world wide. \cite{fond}

The fond usually has someone in charge of the investments called fonds manager. He has the power to invest the money. Additionally to shares fonds also invest in proposals. A proposal is generated by companies where they offer investment opportunities for a specific project or program. The fond manager then bids on the proposals he thinks will be profitable. \cite{proposal}

@@

In May 2016 when the DAO was launched on the ethereum blockchain it was funded with 12.7M ether.
As one ether was worth about 10-12 Dollars at that time, this corresponds to about 150 million dollars and to about 15\% of all the ETH at that time.\cite{understandingDAO}

@@

The split function contains a function called withdrawRewardFor which calls rewardAccount.payOut which in turn calls \textunderscore recipient.call.value. The first two functions are defined in the DAO code and as such can't be changed at will. The call to recipient however is in the attackers hands and as such, he can do there whatever he desires. This function is used to transfer token to an address by calling a function on that address. In code this looks like that: \cite{deconstructingDaoAttack}

@@

The hack was possible through weaknesses in the DAO contract. The ethereum blockchain was not vulnerable and did not aid the attacker. Which is important to note.

@@

After the attacker was done he couldn't instantly retrieve all the funds. He needed to wait 28 days before the proposal to send all the ETH to his address could get through. To somehow retrieve the ether it owuld have to be done in this time. The question was if something can be done. For any kind of fork the network needs to create consensus of the majority within this timeframe. Such a consensus would need a central party which cryptocurrency should not have.

Ethereum is a special case though. Other than Bitcoin where the blockchain didn't start with any bitcoin, the ethereum blockchain strted with an amount of ether wich currently represents 69\% of all ether currently available. \cite{ethSupply} And as such there is such a party, the Ethereum Foundation which stands behind ethereum. \cite{eth} With that the possibility of a fork is possible. The question if it should be done still stays. The following options are possible:

\begin{itemize}
  \item Hard Fork
  \item Soft Fork
  \item Do Nothing
\end{itemize}

\subsection{Hard Fork}

With a hard fork the blockchain could be reverted to before the attack. The DAO contract changed that it contains only a withdraw function and all investors in the DAO would get their money back. For the investors this is probably the best option as they get all their money back without having lost ether. As the DAO contains roughly 15\% of all ether it is said to be too big to fail. \cite{understandingDAO}

However, the blockchain is supposed to be immutable. Many compare the hard fork method to a bailout. A bailout is when a central authority, mostly the state, gives a company, in recent history mostly banks, a lot of money so that they can survive. It was done in the latest financial crisis and sets a dangerous precedent for ethereum as the blockchain would no longer be immutable. A hard fork might be needed if the blockchain itself has a security issue and could be exploited, but as this hack was only exploiting the DAO code it doesn't fall in this category. \cite{understandingDAO}

\subsection{Soft Fork}

With a soft fork the code could be altered in such a way that the DAO and all children of it couldn't be touched. Meaning that any ether invested in those can't be retrieved. This way the attacker had no chance to retrieve his ether and like that he couldn't profit from it. \cite{understandingDAO}

There are many voices against a soft fork too. People argue that ethereum worked as intended and as such it shouldn't be updated. If there are multiple blacklists included into the code it could lead to some unforseen consequences. Also it was only seen as a first step towards the hard fork, which for most blockchain enthusiasts is unthinkable. \cite{understandingDAO}

\subsection{Do Nothing}

A part of the blockchain community argues to do nothing. This way the immutability is not hindered. They argue that investors always take a risk and if one carelessly invests without the necessary precaution, which here would be checking the code for security issues, one assumes unknown risk. They also argue that when this time there is a fork it is a bailout which will lead to future players will demand the same treatment. \cite{understandingDAO}

\subsection{What was done}

The community voted and found that most people were for a change and that the hard fork would be accepted. On July 20, 2016 the hard fork took place. Most miners changed to a new version of the blockchain and kept the name. The miners who were against this change continued on the old chain and continued with the name Ethereum Classic. \cite{fork}

@@

The DAO hack was one of the biggest if not the biggest theft of the history. With a relatively simple attack an unknown entity was able to retrieve value of 45 million dollars. Combined with possible market manipulation he could have gotten a very large amount of money with relatively little effort.

It has shown how important security can be, especially when a lot of money is involved. The hack could have been avoided if more thought would have been invested into the design of the code.

With the decision to create a hard fork, the ethereum foundation has shown that the immutability of the ethereum blockchain can not be viewed as given. They have created precedence where the immutability was less prioritized than the wishes of token holders.
